Manuscript 62524 --- Response to reviewers

### 2025-08-28 Manuscript/Ant_Cactus_RESPONSE.tex
tom's edits on resubmission

\comment{Both reviewers mentioned that the introduction does not quite set up the methods and results of the paper, and I agree, and I want to quickly expand on their comments. 
The introduction is mainly concerned with setting up the question of whether partner diversity in a mutualism can benefit a focal species (via sampling, complementarity, or portfolio effects). 
I agree that this is an interesting question, but I thought that the introduction undersells the results, because the demographic data and IPM offer more than an answer to this one question. 
I think the authors could ‘sell’ their demographic dataset more (just a single intro sentence says “the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available” at Line 101). 
I also thought they could explain in the introduction why even the quantification of the effects of single ant partners across the tree cholla lifecycle, and how ant species turn over on plants, are unique and valuable contributions to the field.}
\response{These are great suggestions -- thank you. We have revised the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data3} and ~\ref{ms-res:long-term-data2}). 
We have also expanded on why the quantification of the complete demographic effects of even single ant partners is a unique and valuable contribution to the field (line ~\ref{ms-res:L111})
}

@@

\comment{I also encourage the authors to pay close attention to responding to several important points raised by Reviewer \#2, especially with regards to 1) what evidence for the portfolio effect would look like in a mutualism in which partners participate sequentially, and 2) providing the reader with a little more ‘hand holding’ when it comes to the rather complex model. Like the reviewer, I cannot claim that I would spot an error in the model if there were one, and it may be helpful to build up to the full Integral Projection Model in more stages, as Reviewer \#2 very helpfully suggests. }
\response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines \ref{ms-res:portfolio_evidence} and Figure \ref{ms-app:portfolio}). 
	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute positive results of this mechanism (lines \ref{ms-res:portfolio_asynchronous}). 
	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{ms-res:portfolio_methods}). }

@@

\comment{Lines 329 and 364: I realize that this is commonplace in some parts of the literature, but I would prefer not to call different simulation scenarios “experiments”}
\response{We have changed the language throughout to "simulations" or "scenarios" rather than "experiments".}

@@

\comment{There is very little mention of some of the supplemental results in the main text (e.g., the herbivory analysis). Please make sure all supplemental results are clearly described in the Results section of the main text.}
\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{res:portfolio_methods}, \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}

@@

\comment{The study "Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism" uses a long-term dataset to look at the demography of a plant with mutualistic ants that provide defense to the plant. 
The estimated parameters and used a stochastic, multi-state IPM to look at the effects of ant mutualists on plant fitness. 
Most significantly, the authors found all ant species increased fitness (finite rate of growth), irrespective of their identity or turnover over the life of the plant. 
Overall, I found the study to be high-quality and novel, and the main finding is somewhat perplexing and would be of interest of the journal's readership. 
Admittedly, I found the framing of the introduction to not quite set up the study strongly and I was worried about the precision of the modeling, but the data and their implications are compelling and interesting. 
The level of precision of the modeling that concerned me was philosophical rather than practical. 
Specifically, I spent a few days thinking about this study with respect to Levins' (1966) strategies of model building in population biology. 
I originally thought that this population model was one that focused on *precision* and either generality or realism, but upon reflection I see this as one that sacrifices precision for generality and realism because the results and interpretation are largely focused on the qualitative outcome of the model, which are the useful and meaningful results. 
I think that it was really just the first paragraph of the introduction that was a sticking point for me given that mutualism "deterioration" and context dependency didn't seem to be important for the study.}
\response{We appreciate the reviewer's positive comments about our manuscript and their thoughtful reflection of how our work relates to broader philosophical considerations of ecological modeling. 
We agree with the Reviewer's conclusion that, in the context of Levins' triangle, our work leans toward generality and realism at the expense of precision. 
That is, we have built a general model of demographic transitions (births, deaths, growth, and partner turnover) that is realistic with respect to our study system, but we query the model for qualitative inference (e.g., asking whether a multi-partner scenario yields higher or lower fitness than the single best partner) and not quantitative precision. 
We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}). ******** Tom have you already done this? Or is this something you would like me to do?}

@@

\comment{L. 22: I feel like this paragraph doesn't add much and that if you nix everything after the first sentence and open with what's in paragraph 2, that would be more suitable for this study. 
It also frees some room to discuss generalized/diffuse mutualisms further.}
\response{Thank you for the helpful suggestion. 
We have streamlined the introduction paragraph by removing extraneous details after the first sentence and opened the section with the content previously in paragraph 2. 
This revision improves the focus of the introduction on the core aspects of our study. 
Additionally, we have expanded the discussion on generalized/diffuse mutualisms}, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).

@@

\comment{Fig. 5, caption: ". . . based on [plant] size".}
\response{We have revised the caption of Figure 5 to clarify what is meant by "based on plant size" and to explicitly describe the relationship being visualized (Figure~\ref{ms-fig:Ant_Transition}).}

@@

\comment{I would like to know what evidence for the portfolio effect would look like, so that I can carry that through the methods and see that this is what you are measuring, and into the results, so I can compare it to what you actually found.}
\response{Great suggestion. 
	We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_theory}).
	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{ms-res:portfolio_criteria}).}}

@@

\comment{27: Compared to which interactions ("such as…")? I think this intro material obstructs you from rolling out the background we really need.}
\response{We have revised the first paragraph to reduce unnecessary and/or uninformative information to speed the reader into what is needed to understand the rest of the paper (line \ref{ms-res:L27}).}

>>>>

\comment{Both reviewers mentioned that the introduction does not quite set up the methods and results of the paper, and I agree, and I want to quickly expand on their comments. 
The introduction is mainly concerned with setting up the question of whether partner diversity in a mutualism can benefit a focal species (via sampling, complementarity, or portfolio effects). 
I agree that this is an interesting question, but I thought that the introduction undersells the results, because the demographic data and IPM offer more than an answer to this one question. 
I think the authors could ‘sell’ their demographic dataset more (just a single intro sentence says “the long-term data needed to quantify fluctuations in ant-plant interactions are rarely available” at Line 101). 
I also thought they could explain in the introduction why even the quantification of the effects of single ant partners across the tree cholla lifecycle, and how ant species turn over on plants, are unique and valuable contributions to the field.}
\response{These are great suggestions -- thank you. We have revised the introduction to better reflect the importance of long term data (see lines ~\ref{ms-res:long-term-data}, ~\ref{ms-res:long-term-data2}). 
We have also expanded on why the quantification of the complete demographic effects of even single ant partners is a unique and valuable contribution to the field (line ~\ref{ms-res:L111})
}

@@

\comment{I also encourage the authors to pay close attention to responding to several important points raised by Reviewer \#2, especially with regards to 1) what evidence for the portfolio effect would look like in a mutualism in which partners participate sequentially, and 2) providing the reader with a little more ‘hand holding’ when it comes to the rather complex model. Like the reviewer, I cannot claim that I would spot an error in the model if there were one, and it may be helpful to build up to the full Integral Projection Model in more stages, as Reviewer \#2 very helpfully suggests. }
\response{We agree about the importance of Reviewer 2's comments, to which we respond in greater detail below. 
	In brief, we have added new analyses, as suggested, demonstrating a `positive test case' where our model exhibits positive effects of partner diversity through the portfolio effect (lines \ref{ms-res:portfolio_evidence} and Figure \ref{ms-app:portfolio}). 
	We have also edited the introduction to better set up readers for understanding the portfolio effect and what would constitute evidence for this mechanism (lines \ref{ms-res:portfolio_asynchronous}). 
	Finally, we have thoroughly revised our presentation of the model so that it develops in stages, which we hope will make it clearer and more digestible (lines \ref{ms-res:portfolio_methods}). }

@@

\comment{Lines 329 and 364: I realize that this is commonplace in some parts of the literature, but I would prefer not to call different simulation scenarios “experiments”}
\response{We have changed the language throughout to ``simulations'' or ``scenarios'' rather than ``experiments''.}

@@

\comment{There is very little mention of some of the supplemental results in the main text (e.g., the herbivory analysis). Please make sure all supplemental results are clearly described in the Results section of the main text.}
\response{Thank you for pointing this out. We now provide clearer connections between the main text and that additional information and results provided in the supplement (lines \ref{ms-res:appendixc}), \ref{ms-res:portfolio_results}, \ref{ms-res:seeds}, \ref{ms-res:herbivory}.}

@@

\comment{The study "Demographic consequences of partner diversity and turnover in a multi-species ant-plant mutualism" uses a long-term dataset to look at the demography of a plant with mutualistic ants that provide defense to the plant. 
The estimated parameters and used a stochastic, multi-state IPM to look at the effects of ant mutualists on plant fitness. 
Most significantly, the authors found all ant species increased fitness (finite rate of growth), irrespective of their identity or turnover over the life of the plant. 
Overall, I found the study to be high-quality and novel, and the main finding is somewhat perplexing and would be of interest of the journal's readership. 
Admittedly, I found the framing of the introduction to not quite set up the study strongly and I was worried about the precision of the modeling, but the data and their implications are compelling and interesting. 
The level of precision of the modeling that concerned me was philosophical rather than practical. 
Specifically, I spent a few days thinking about this study with respect to Levins' (1966) strategies of model building in population biology. 
I originally thought that this population model was one that focused on *precision* and either generality or realism, but upon reflection I see this as one that sacrifices precision for generality and realism because the results and interpretation are largely focused on the qualitative outcome of the model, which are the useful and meaningful results. 
I think that it was really just the first paragraph of the introduction that was a sticking point for me given that mutualism "deterioration" and context dependency didn't seem to be important for the study.}
\response{We appreciate the reviewer's positive comments about our manuscript and their thoughtful reflection of how our work relates to broader philosophical considerations of ecological modeling. 
We agree with the Reviewer's conclusion that, in the context of Levins' triangle, our work leans toward generality and realism at the expense of precision. 
That is, we have built a general model of demographic transitions (births, deaths, growth, and partner turnover) that is realistic with respect to our study system, but we query the model for qualitative inference (e.g., asking whether a multi-partner scenario yields higher or lower fitness than the single best partner) and not quantitative precision. 
We now use stronger ``signposting'' in the presentation of the model to make this clearer for readers (e.g., lines \ref{}, \ref{}). 
%******** Tom have you already done this? Or is this something you would like me to do?
}

@@

\comment{L. 22: I feel like this paragraph doesn't add much and that if you nix everything after the first sentence and open with what's in paragraph 2, that would be more suitable for this study. 
It also frees some room to discuss generalized/diffuse mutualisms further.}
\response{Thank you for the helpful suggestion. 
We have streamlined the introduction paragraph by removing extraneous details after the first sentence. We now open with the content previously in paragraph 2.} 
%Additionally, we have expanded the discussion on generalized/diffuse mutualisms, as recommended, providing clearer context for the ecological significance of our work (line~\ref{ms-res:L22}).

@@

\comment{Fig. 5, caption: ". . . based on [plant] size".}
\response{We have revised the caption of Figure 5 to clarify what is meant by ``based on plant size'' and to explicitly describe the relationship being visualized (Figure~\ref{ms-fig:Ant_Transition}).}

@@

\comment{I would like to know what evidence for the portfolio effect would look like, so that I can carry that through the methods and see that this is what you are measuring, and into the results, so I can compare it to what you actually found.}
\response{Great suggestion. 
	We have revised the Introduction to include a more detailed conceptual explanation of how the portfolio effect can operate in mutualisms with sequential partner turnover. 
	Specifically, we now define the portfolio effect as occurring when inter-annual variation in partner identity stabilizes plant fitness across years, due to asynchronous responses of different ant partners to environmental variability (line~\ref{ms-res:portfolio_asynchronous}).
	As requested, we now provide criteria for what evidence would be required to conclude that the portfolio effect contributes to benefits of partner diversity (line \ref{ms-res:portfolio_criteria}).}

@@

\comment{27: Compared to which interactions ("such as…")? I think this intro material obstructs you from rolling out the background we really need.}
\response{We have heavily revised the first paragraph as described above.}
